\section{Using pbdPAPI}
\label{sec:use}


The \thispackage package has 2 distinct interfaces for gathering performance 
data, one low-level and ``one'' high-level.  The low-level interface is 
reminiscent of \PAPI's own native \C interface, and ``the'' high-level interface 
is a collection of similar-looking wrappers around this interface.

Note that not all utilities are available on all hardware platforms.  You can 
check which profiler events are available on your hardware platform using the 
\thispackage utility \code{papi.avail()} called with no argument.


\subsection{High-Level Interface}

We offer numerous useful simplified interfaces for profiling \R code, shown in 
the table below:
\begin{center}
\vspace{0.2cm}
\begin{tabular}{ll} \hline\hline
Function & Description of Measurement \\ \hline
\code{system.cache()} & Cache misses, hits, accesses, and reads \\
\code{system.cpuormem()} & Characterize code as CPU or RAM bound (see docs 
for definition) \\
\code{system.epc()} & Events per cycle \\
\code{system.flips()} & Time, floating point instructions, and Mflips \\
\code{system.flops()} & Time, floating point operations, and Mflops \\
\code{system.idle()} & Idle cycles \\
\code{system.utilization()} & CPU utilization (see documentation for 
definition) \\
\hline\hline
\end{tabular}
\vspace{0.2cm}
\end{center}

Each of these utilities behaves somewhat like \R's own \code{system.time()} 
function.  Simply pass a valid \R expression in for the \code{expr} argument 
for any of these functions and the appropriate measurement will be returned.

See \secref{sec:examples} and/or the package demos for some example usage.


\subsection{Low-Level Interface}

The low-level interface is provided for users who more familiar with hardware and may have 
questions about the operation of a block of code that do not easily fall into the the set of 
wrappers that make up the high-level interface.  Even this comes in two forms.  The first is the 
\code{system.event()} interface, which looks like the high-level interface functions, except that 
you must explicitly supply a \PAPI events vector, with event names stored as strings.  For a list 
of all possible events, and how to determine which are supported on your platform, see the \R help 
\code{?papi.avail}.

As an example, suppose you wanted to measure the level 1 data cache misses and L1 data cache hits. 
After looking up the names for these events in the help file mentioned just above, you would simply 
call
\begin{lstlisting}[language=rr]
events <- c("PAPI_L1_DCM", "PAPI_L1_DCH")
system.event(1+1, events=events)
\end{lstlisting}

Here, replace \code{1+1} with your preferred computation.

The final form of the low-level interface is essentially a deconstruction of the code making up the 
body of the \code{system.event()} function.  This would allow you to skip several forms of error 
checking (if you want to do that for some reason).  Primarily its purpose is to resemble the 
native \C-level \PAPI interface.  We really don't recommend you use this, but it's there if you 
really believe you need it.

Revisiting the above example, you would call:
\begin{lstlisting}[language=rr]
events <- c("PAPI_L1_DCM", "PAPI_L1_DCH")

papi.start(events=events)
1+1
papi.stop(events=events)
\end{lstlisting}


\